---
output:
  word_document: default
  html_document: default
  pdf_document: default
---
# Reproducable Research, Project 1

## Part 1: Loading and preprocessing the data

Data found here: https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip
Download data and unzip files into your working direcory
Next, Load the data into R:

```{r opendoc, echo=TRUE}
setwd("C:/Users/LindseyBehrens/OneDrive - Genective/Documents/R Working/Coursera")

Data <- read.csv("activity.csv")
head(Data)
```


Take a look at the Data:
```{r lookatdata, echo=TRUE}
str(Data)
```



Notice that the 'date' column is read as a character instead of a date. Let's fix that:

```{r makedate, echo=TRUE}
Data$date <- as.Date(Data$date)
str(Data)
```







## Part 2: What is mean total number of steps taken per day?

We will ignore the missing values here. 
Lets fine the total number of steps:
```{r totalSteps, echo=TRUE}
totalSteps <- sum(Data$steps, na.rm = TRUE)
totalSteps
```

570608 total steps!



Lets look at the distribution in a histogram:
```{r HistogramSteps, echo=TRUE}
library(ggplot2)
Histogram <- ggplot(Data, aes(x=steps)) +
    geom_histogram(binwidth=25)
Histogram
```

That's a lot of 0 steps.



Lets look at the mean and median number of steps:

```{r meanSteps, echo=TRUE}
meanSteps <- mean(Data$steps, na.rm = TRUE)
meanSteps
medianSteps <- median(Data$steps, na.rm = TRUE)
medianSteps
```

Using this code we found that the mean number of steps was 37.3826 and the median was 0 steps. 







## Part 3: What is the average daily activity pattern?

Lets make a graph to look at the average number of steps in each 5-minute interval of the day
```{r intervals, echo=TRUE}
library(dplyr)
Averaged <- Data %>%
  group_by(interval) %>%
  summarize(steps = mean(steps, na.rm = TRUE))


AverageSteps <- ggplot(Averaged, aes(y=steps, x= interval)) +
  geom_line()
AverageSteps
```


Now lets find the interval with the highest average number of setps

```{r highestinterval, echo=TRUE}
Highest <- seq(along=Averaged$steps)[Averaged$steps==max(Averaged$steps)]
Averaged[Highest,]
```
Interval 835 with an average of 206 steps!






## Part 4: Imputing missing values

Note that there are a number of days/intervals where there are missing values (coded as NA). 
The presence of missing days may introduce bias into some calculations or summaries of the data.

Find the total number of rows with missing values:
```{r detect NA, echo=TRUE}
Data_No_NA <- Data[complete.cases(Data), ]
NA_count <- nrow(Data) - nrow(Data_No_NA)
NA_count
```
There are 2304 missing values


Now we will fill the missing NA's with the median
```{r replaceNA, echo=TRUE}
library(dplyr)
library(tidyr)
Imputed_Data <- Data %>% 
  mutate(steps = replace_na(steps,median(steps, na.rm = TRUE)))
nrow(Imputed_Data) ==  nrow(Data)
head(Imputed_Data)
```
After replacing the NA's with the median the data sets are the same length and there are no more NA's


Now Lets make a histogram of the total number of steps taken each day
```{r Histogram2, echo=TRUE}
Histogram2 <- ggplot(Imputed_Data, aes(x=steps)) +
  geom_histogram(binwidth=25)
Histogram2
```

And look at the new mean and median
```{r means2, echo=TRUE}
meanSteps2 <- mean(Imputed_Data$steps, na.rm = TRUE)
meanSteps2
medianSteps2 <- median(Imputed_Data$steps, na.rm = TRUE)
medianSteps2
```

Do these values differ from the estimates from the first part of the assignment? 
The mean is slightly lower and the median is the same. 
```{r comparemeans, echo=TRUE}
meanSteps
meanSteps2
medianSteps
medianSteps2
```


What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r total2, echo=TRUE}
totalSteps2 <- sum(Imputed_Data$steps, na.rm = TRUE)
totalSteps2
totalSteps
```
The same! The median was 0. 








## Part 5: Are there differences in activity patterns between weekdays and weekends?


Create a new factor variable in the dataset with two levels indicating whether a given date is a weekday or weekend day.
```{r weekdayClass, echo=TRUE}
Imputed_Data$Weekday <- weekdays(Imputed_Data$date)
Day <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
End <- c("Saturday", "Sunday")

Imputed_Data$Weekend <-  Imputed_Data$Weekday %in% End
Imputed_Data <- Imputed_Data %>%
  mutate(DayClass = case_when(
    Weekend == TRUE ~ "Weekend",
    Weekend == FALSE ~ "Weekday"
  ))

head(Imputed_Data)
```


Make a time series plot of the 5-minute interval and the average number of steps taken, 
averaged across all weekday days or weekend days
```{r weekdayGraph, echo=TRUE}
Classes <- unique(Imputed_Data$DayClass)
Averaged2 <- data.frame(matrix(ncol=3, nrow=0))
for (Class in Classes) {
  set <- subset(Imputed_Data, DayClass == Class)
  Average <- set %>%
    group_by(interval) %>%
    summarize(steps = mean(steps, na.rm = TRUE))
  Average$DayClass <- Class
  Averaged2 <- rbind(Averaged2, Average)
}

AverageSteps2 <- ggplot(Averaged2, aes(y=steps, x= interval)) +
  geom_line() +
  facet_grid( ~ DayClass)
AverageSteps2
```




















